The Precambrian paleogeography of Laurentia

### 2021-07-23 Manuscript/Laurentia_Paleogeo_Manuscript.tex
manuscript edits associated with proof corrections

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{../Figures/Fig8_Rodinia_Reconstruction.pdf}
\caption{\textbf{Paleogeographic reconstructions of Laurentia and other select conjugate Proterozoic continents leading up to Rodinia assembly in the late Mesoproterozoic and to its initial break-up in the Neoproterozoic}. The hypothesized connection between Siberia and Laurentia is implemented following \cite{Evans2016b} who interpret this relationship as persistent from 1.7 to 0.7 Ga. The reconstruction of North China to Laurentia follows \cite{Ding2021a}. The reconstruction implements Kalahari and Amazonia cratons as conjugates with Laurentia in the Grenvillian orogeny \citep{Hoffman1991a}. The Australia-East Antarctica relationship with Laurentia follows \cite{Swanson-Hysell2012a} and is similar to the Neoproterozoic reconstruction between the continents of \cite{Li2011a} and implementing their relative rotation of North Australia relative to South and West Australia. This configuration back to ca. 1140 Ma is consistent with a comparison between the Laurentia poles of that age and the coeval Mt. Isa dikes pole from North Australia and with the Keweenawan Track if the the Nonesuch and Freda poles are interpreted to be ca. 1080 (consistent with chronostratigraphic constraints; \citealp{Slotznick2018b}) with further motion by ca. 1070 Ma. The time slices show the rapid motion of Laurentia implied by the paleomagnetic poles which is consistent with the timing of collisional orogenesis associated with the Grenvillian orogeny. The assembled Rodinia persisted until initial rifting ca. 775 Ma with episodic rifting continuing until ca. 530 Ma.}
\label{fig:Grenville_reconstructions}
\end{figure*}

@@

%%%%%%%%%%%%% Acknowledgements %%%%%%%%%%%%%
\footnotesize
\subsection*{Acknowledgements}
This work was supported by NSF CAREER Grant EAR-1847277. The manuscript benefited from reviews from Athena Eyster, David Evans, and Lauri Pesonen as well as discussions and manuscript feedback from Francis Macdonald, Yuem Park, Toby Rivers, Sarah Slotznick, Justin Strauss, and Yiming Zhang. Bruce Buffett provided insights on the implications of a stagnant lid regime for the generation of the geomagnetic field. Many participants in the Nordic Paleomagnetism Workshop have contributed to the compilation and evaluation of the pole list utilized herein. Particular acknowledgement goes to David Evans for maintaining and distributing the compiled pole lists. GPlates, and in particular the pyGPlates API, was utilized in this work \citep{Muller2018b}. Figures were made using Matplotlib \citep{Hunter2007a} in conjunction with cartopy \citep{Met-Office2010a} and pmagpy \citep{Tauxe2016a} within an interactive Python environment \citep{Perez2007a}. The chapter text as well as code, data, and reconstructions used in this paper are openly available and licensed for any form of reuse with attribution (CC BY 4.0) in this repository: \url{https://github.com/Swanson-Hysell-Group/Laurentia_Paleogeography} which is also archived on Zenodo.

>>>>

\begin{figure*}
\centering
\includegraphics[width=\textwidth]{../Figures/Fig8_Rodinia_Reconstruction.pdf}
\caption{\textbf{Paleogeographic reconstructions of Laurentia and other select conjugate Proterozoic continents leading up to Rodinia assembly in the late Mesoproterozoic and to its initial break-up in the Neoproterozoic}. The hypothesized connection between Siberia and Laurentia is implemented following \cite{Evans2016b} who interpret this relationship as persistent from 1.7 to 0.7 Ga. The reconstruction of North China to Laurentia follows \cite{Ding2021a}. The omission of South China from Rodinia follows ​\cite{Park2021b}​. The reconstruction implements Kalahari and Amazonia cratons as conjugates with Laurentia in the Grenvillian orogeny \citep{Hoffman1991a}. The Australia-East Antarctica relationship with Laurentia follows \cite{Swanson-Hysell2012a} and is similar to the Neoproterozoic reconstruction between the continents of \cite{Li2011a} and implementing their relative rotation of North Australia relative to South and West Australia. This configuration back to ca. 1140 Ma is consistent with a comparison between the Laurentia poles of that age and the coeval Mt. Isa dikes pole from North Australia and with the Keweenawan Track if the the Nonesuch and Freda poles are interpreted to be ca. 1080 (consistent with chronostratigraphic constraints; \citealp{Slotznick2018b}) with further motion by ca. 1070 Ma. The time slices show the rapid motion of Laurentia implied by the paleomagnetic poles which is consistent with the timing of collisional orogenesis associated with the Grenvillian orogeny. The assembled Rodinia persisted until initial rifting ca. 775 Ma with episodic rifting continuing until ca. 530 Ma.}
\label{fig:Grenville_reconstructions}
\end{figure*}

@@

%%%%%%%%%%%%% Acknowledgements %%%%%%%%%%%%%
\footnotesize
\subsection*{Acknowledgements}
This work was supported by NSF CAREER Grant EAR-1847277. The manuscript benefited from reviews from Athena Eyster, David Evans, and Lauri Pesonen as well as discussions and manuscript feedback from Francis Macdonald, Yuem Park, Toby Rivers, Sarah Slotznick, Justin Strauss, and Yiming Zhang. Bruce Buffett provided insights on the implications of a stagnant lid regime for the generation of the geomagnetic field. Many participants in the Nordic Paleomagnetism Workshop have contributed to the compilation and evaluation of the pole list utilized herein. Particular acknowledgement goes to David Evans for maintaining and distributing the compiled pole lists. GPlates, and in particular the pyGPlates API, was utilized in this work \citep{Muller2018b}. Figures were made using Matplotlib \citep{Hunter2007a} in conjunction with cartopy \citep{Met-Office2010a} and pmagpy \citep{Tauxe2016a} within an interactive Python environment \citep{Perez2007a}. The chapter text as well as code, data, and reconstructions used in this paper are openly available and licensed for any form of reuse with attribution (CC BY 4.0) in this repository: \url{https://github.com/Swanson-Hysell-Group/Laurentia_Paleogeography} which is also archived on Zenodo (https://doi.org/10.5281/zenodo.5129624).
